OpenMPIR

### 2017-09-06 paper.tex
Initial comments from WM

\documentclass[sigconf]{acmart}

@@

Internal representations (IRs) are the compiler objects on which almost all
analyses and optimizations are performed. Historically, a major drawback for
parallel codes is that IRs don't have the ability to represent, and therefore
reason about, parallelism. This prevents the application of common
optimizations to concurrent code. The recent work of Schardl et al.
\cite{shardl2017} has shown that real world compiler IRs can be extended to
represent fork-join parallelism. By extending the LLVM instruction set with
three instructions, they were able to capture all of the semantics to implement
the semantics of Cilk, an extension to C. They suggested that a similar
approach could be taken with OpenMP tasks.

@@

LLVM is arguably the most successful IR to date \cite{llvm}. By implementing a simple
single static assignment (SSA) representation, it enables powerful optimizations 
that are effective for a wide range of programming languages and hardware
\cite{llvm imps}.

@@

Addressing these concerns this is a major motivation for our turning to Tapir.
By standardising a powerful but simple internal representation and building
frontends and backends for it, one both make it easier to build optimizations
and analyses for a given programming model, and avoid duplicating that work
across different implementations. We'll return to this issue later in the paper
when discussing potential limitations and challenges for Tapir for other
hardware and

@@

The surprising fact that program behaviour wasn't changed significantly is
worth visiting here. We surmise that it is likely due to the variable semantics
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics.  In particular, \emph{OpenMP doesn't specify any memory model for
shared variables}. This results in the Cilk style of only writing to a shared
variable at the end of a parallel section as a perfectly valid implementation
of OpenMP's specification. In contrast, existing OpenMP implementations often
write to shared variables on every write. This is a potential for performance
discrepancy due to differing implementation-defined behaviour, and we will 
return to it when discussing performance on the Barcelona OpenMP task suite.
There is of course the fact that technically, by copying the value of variables
back to the surrounding context even for \texttt{private} variables, one isn't
following the specification. We surmise that in practice this hasn't had an
effect due to the fact that \texttt{private} variables are generally used for
performance, rather than their semantic properties.

>>>>

\documentclass[sigconf]{acmart}
\newcommand{\wmnote}[1]{{\scriptsize \color{red} [[ Billy: #1]]}}

@@

Internal representations (IRs) are the \wmnote{the phrase compiler objects seems a bit odd to me}compiler objects on which almost all
analyses and optimizations are performed. Historically, a major drawback for
parallel codes is that IRs don't have the ability to represent, and therefore
reason about, parallelism. This prevents the application of common
optimizations to concurrent code. The recent work of Schardl et al.
\cite{shardl2017} has shown that real world compiler IRs can be extended to
represent fork-join parallelism. By extending the LLVM instruction set with
three instructions, they were able to capture all of the semantics to implement
the semantics of Cilk, an extension to C. They suggested that a similar
approach could be taken with OpenMP tasks.

@@

\wmnote{This feels like we're trying to convince someone about LLVM: perhaps LLVM has gained traction in the academic community for both its simplicity and relatively painless modification}LLVM is arguably the most successful IR to date \cite{llvm}. By implementing a simple
single static assignment (SSA) representation, it enables powerful optimizations 
that are effective for a wide range of programming languages and hardware
\cite{llvm imps}.

@@

Addressing these concerns this is a major motivation for our \wmnote{use of?}turning to Tapir.
\wmnote{this sentence feels very loose/run-on} By standardising a powerful but simple internal representation and building
frontends and backends for it, one both makes it easier to build optimizations
and analyses for a given programming model, and avoids duplicating work
across different implementations. We'll return to this issue later in the paper
when discussing potential limitations and challenges for Tapir for other
hardware and

@@

The surprising fact that program behaviour wasn't changed significantly is
worth \wmnote{discussion?}visiting here. We surmise that it is likely due to the variable semantics
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics.  In particular, \emph{OpenMP doesn't specify any memory model for
shared variables}. This results in the Cilk style of only writing to a shared
variable at the end of a parallel section as a perfectly valid implementation
of OpenMP's specification. In contrast, existing OpenMP implementations often
write to shared variables on every write. This is a potential for performance
discrepancy due to differing implementation-defined behaviour, and we will 
return to it when discussing performance on the Barcelona OpenMP task suite.
There is of course the fact that technically, by copying the value of variables
back to the surrounding context even for \texttt{private} variables, one isn't
following the specification. We surmise that in practice this hasn't had an
effect due to the fact that \texttt{private} variables are generally used for
performance, rather than their semantic properties.
